Detecting Hate Speech with GPT-3

### 2022-03-16 outputs/documents/paper.Rmd
Final read of paper

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    toc: false
    fig_caption: true
    latex_engine: xelatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
  -  \usepackage{amsmath}
biblio-style: apalike
title: "Detecting Hate Speech with GPT-3"
thanks: "Code and data are available at: https://github.com/kelichiu/GPT3-hate-speech-detection. We gratefully acknowledge the support of Gillian Hadfield, the Schwartz Reisman Institute for Technology and Society, and OpenAI for providing access to GPT-3 under the academic access program. We thank two anonymous reviews and the editor, as well as Amy Farrow, Haoluan Chen, John Giorgi, Mauricio Vargas Sepúlveda, Monica Alexander, Noam Kolt, and Tom Davidson for helpful discussions and suggestions. Please note that we have added asterisks to racial slurs and other offensive content in this paper, however the inputs and outputs did not have these.  Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Ke-Li Chiu
  affiliation: University of Toronto
- name: Annie Collins
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: University of Toronto and Schwartz Reisman Institute
abstract: "Sophisticated language models such as OpenAI's GPT-3 can generate hateful text that targets marginalized groups. Given this capacity, we are interested in whether large language models can be used to identify hate speech and classify text as sexist or racist. We use GPT-3 to identify sexist and racist text passages with zero-, one-, and few-shot learning. We find that with zero- and one-shot learning, GPT-3 can identify sexist or racist text with an accuracy between 57 per cent and 68 per cent depending on the category of text and type of learning. With few-shot learning, the model's accuracy can be as high as 88 per cent. Large language models have a role to play in hate speech detection, and with further development could eventually be used to counter hate speech."
keywords: "GPT-3; natural language processing; quantitative analysis; hate speech."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
endnote: no
graphics: yes
---

@@

The text produced by sophisticated NLP models can be hateful. In particular, there have been many examples of text being generated that targets marginalized groups based on their sex, race, sexual orientation, and other characteristics. For instance, 'Tay' was a chatbot that was released by Microsoft in 2016 on Twitter. Within hours of being released, some of its tweets were sexist. Large language models are trained on enormous datasets from various sources. This means they usually contain untruthful statements, human biases, and abusive language. Further, even when the models do not possess intent, they risk producing text that is offensive or discriminatory, and thus cause unpleasant, or even triggering, interactions [@bender2021dangers].

Often the datasets that underpin these models consist of, essentially, the whole public internet. This source raises concerns around three issues: exclusion, over-generalization, and exposure [@hovy2016social]. Exclusion happens due to the demographic bias in the dataset. In the case of language models that are trained on US and UK English scraped from the Internet, datasets may be disproportionately white, male, and young. Therefore, it is not surprising to see white supremacist, misogynistic, and ageist content being over-represented in training datasets [@bender2021dangers]. Over-generalization stems from the assumption that what we see in the dataset represents what occurs in the real world. Words such as 'always', 'never', 'everybody', or 'nobody' are frequently used for rhetorical purpose instead of their literal meanings. However, language models do not always recognize this and make inference based on generalized statements using these words. For instance, hate speech commonly uses generalized language for targeting a group such as 'all' and 'every', and a model trained on these statements may generate similarly generalized and harmful statements. Finally, exposure refers to the relative attention, and hence consideration of importance, given to something. In the context of NLP this may be reflected in the emphasis on English-language created under particular circumstances, rather than another language or circumstances that may be more prevalent.

@@

Even though large NLP models do not have a real understanding of language, the vocabularies and the construction patterns of harmful languages can be thought of as known to them. We show that this knowledge can be used to identify abusive language and even hate speech. We consider 120 different extracts that have been categorized as 'racist', 'sexist', or 'neither' in single-category settings (zero-shot, one-shot, and few-shot) and 243 different extracts in mixed-category few-shot settings. We ask GPT-3 to classify these based on zero-, one-, and few-shot learning, with and without instruction. We find that the model performs best with mixed-category few-shot learning. In that setting the model can accurately classify around 86 per cent of the racist extracts and 88 per cent of sexist extracts. If language models can be used to identify abusive language, then not only is there potential for them to counter the production of abusive language by humans, but they could also potentially self-police.

@@

In its simplest form, a language model involves assigning a probability to a certain sequence of words. For instance, the sequence 'the cat in the hat' is probably more likely than 'the cat in the computer'. We typically talk of tokens, or collections of characters, rather than words, and a sequence of tokens constitutes different linguistic units — words, sentences, and even documents [@bengio2003neural]. Language models predict the next token based on inputs. If we consider each token in a vocabulary as a dimension, then the dimensionality of language quickly becomes large [@rosenfeld2000two]. Over time a variety of statistical language models have been created to nonetheless enable prediction. The n-gram is one of the earliest language models. It works by considering the co-occurrence of tokens in a sequence. For instance, given the four-word sequence, 'the cat in the', it is more likely that the fifth word is 'hat' rather than 'computer'. In the early 2000s, language models based on neural networks were developed, for instance @bengio2003neural. These were then built on by word embeddings language models in the 2010s in which the distance between tokens represents how related those tokens are, for instance @turian2010word. In 2017, @vaswani2017attention introduced the Transformer, which marked a new era for language models. The Transformer is a network architecture for neural networks that can be trained more quickly than many other approaches [@vaswani2017attention]. Now most representative pre-trained language models, such as Google's: Bidirectional Encoder Representations from Transformers (BERT) [@devlin2018bert], as well as OpenAI's: Generative Pre-trained Transformer (GPT)-2 [@radford2019language], and GPT-3 [@brown2020language], are built on this architecture. These models are widely used, for instance BERT is used by Google search.

@@

There is no commonly held definition of hate speech. Different legal jurisdictions have different definitions, as do different companies and other groups. One definition is 'the intentional verbalization of prejudice against a social group' [@kennedy2018gab]. Detecting hate speech is difficult because the definition of hate speech varies depending on a complex intersection of the topic of the assertion, the context, the timing of the post, synchronized world events, and the identity of speaker and recipient [@schmidt2017survey]. Moreover, it is difficult to distinguish hate speech from offensive language [@davidson2017automated]. Hate speech detection is of interest to academic researchers in a variety of domains including computer science [@addressinghatespeech] and sociology [@davidson2017automated]. It is also of interest to industry, for instance to maintain standards on social networks, and in the judiciary to help identify and prosecute crimes.  Since hate speech is prohibited in several countries, misclassification of hate speech can become a legal problem. For instance, in Canada, speech that contains 'public incitement of hatred' or 'wilful promotion of hatred' is specified by the Criminal Code [@act2021justice]. Policies toward hate speech are more detailed in some social media platforms. For instance, the Twitter Hateful Conduct Policy states:

@@

There has been a large amount of research focused on detecting hate speech. As part of this process, various hate speech datasets have been created and examined. For instance, @waseem2016hateful create a dataset that captures hate speech in the form of racist and sexist language that includes domain expert annotation. They use Twitter data, and annotate 16,914 tweets: 3,383 as sexist, 1,972 as racist, and 11,559 as neither. They had a high degree of annotator agreement. Most of the disagreements were to do with sexism, and often explained by an annotator lacking apparent context. @davidson2017automated trains a classifier to distinguish between hate speech and offensive language. To define hate speech, they use an online 'hate speech lexicon containing words and phrases identified by internet users as hate speech'. Even these datasets have bias. For instance, @davidson2019racial found racial bias in five different sets of Twitter data annotated for hate speech and abusive language. They found that tweets written in African American English are more likely to be labeled as abusive.

@@

We use the ETHOS dataset created by @mollas2020ethos. ETHOS is based on comments found in YouTube and Reddit. The ETHOS YouTube data is collected through Hatebusters [@anagnostou2018hatebusters]. Hatebusters is a platform that collects comments from YouTube and assigns a 'hate' score to them using a support vector machine. That hate-score is only used to decide whether to consider the comment further or not. The Reddit data is collected from the Public Reddit Data Repository [@baumgartner2020pushshift]. The classification is done by contributors to a crowd-sourcing platform. They are first asked whether an example contains hate speech, and then, if it does, whether it incites violence and other additional details. The dataset has two variants: binary and multi-label. In the binary dataset comments are classified as hate or non-hate based. In the multi-label variant comments are evaluated on measures that include violence, gender, race, ability, religion, and sexual orientation. The dataset that we use is from the ETHOS dataset and hence contain typos, misspelling, and offensive content.

We use all of the 998 statements in the ETHOS dataset that have a binary classification of hate speech or not hate speech. Of these, the 433 statements that contain hate speech additionally have labels that classify the content. For instance, does the comment have to do with violence, gender, race, nationality, disability, etc. We considered all of the 136 statements that contain race-based hate speech, and we focus on the 76 whose race-based score is at least 0.5, meaning that at least 50 per cent of annotators agreed. Similarly, we considered all of the 174 statements that contain gender-based hate speech, and again focused on the 84 whose gender-based score is at least 0.5. To create a balanced dataset, we select 120 of these statements to result in 60 sexist statements and 60 racist statements. For comparison purposes, we randomly draw 120 statements that are classified as not hate speech. The balanced sample enables us to more easily understand our results.

@@

Under the few-shot setting, three examples are provided to the model. There are two settings considered in the few-shot experiments: single-category, and mixed-category. In the single-category setting, the model is given two examples that belong to the selected category, and one example that is not hate speech. The answer is expected to be binary. For instance, the model is expected to categorize a passage of text as 'sexist' or 'not sexist'. An example of this is:

@@

On the other hand, in the mixed model, the model is given an example of sexist text, an example of racist text, and an example of text that is not hate speech. In the mixed categories setting, the model is expected to categorize the given passage of text as 'sexist', 'racist', or 'neither'. To simplify the classification procedure and allow for comparisons between different example inputs, we created ten sets of examples against which 240 comments were each classified. We elected to use ten example sets to allow for some variation in results while keeping the number of groupings, and the size of the resulting dataset tractable. The example sets each consist of one sexist, one racist, and one 'neither' comment and were created using the comments at indexes one through ten from each dataset. Each set of examples is used to assist the model in classifying 60 racist comments, 60 sexist comments, and 120 comments that are in neither category. The inputted comments were the first _n_ + 1 comments from their respective datasets with the _i_^th^ comment dropped corresponding to the use of example set _i_. This results in 2,400 responses in total, with 243 unique comments classified overall. The ten sets of examples are shown in Appendix \@ref(appendxa). An example of this is:

@@

We assess GPT-3's performance in all settings using accuracy, precision, recall, and F1 score. Accuracy is the proportion of correctly classified comments (hate speech and non-hate speech) out of all comments classified. Precision is the proportion of hate speech comments correctly classified out of all comments classified as hate speech (both correctly and incorrectly). Recall is the proportion of hate speech comments correctly classified out of all hate speech comments in the dataset (both correctly and incorrectly classified). The F1 score is the harmonic mean of precision and recall. In the case of hate speech classification, it is better to have a model with high recall, meaning a model that can identify a relatively high proportion of the hate speech text within a dataset. But the F1 score can provide a more well-rounded metric for model performance and comparison.

@@

The overall results of the zero-shot experiments are presented in Table \@ref(tab:zeroshot-summary). Tables \@ref(tab:zeroshot-racism), \@ref(tab:zeroshot-sexism), and \@ref(tab:zeroshot-hate), in Appendix \@ref(appendixbzeroshot) contain additional detail. The model has 34 matches (true positives and negatives) and 26 mismatches (false positives and negatives) in the sexist category, and 41 matches and 19 mismatches in the racist category. The model performs more accurately when identifying racist comments, with an accuracy of 68 per cent, compared with identifying sexist comments, with an accuracy of 57 per cent, however the F1 score for each type of hate speech is approximately the same. The overall ratio of matches and mismatches is 75:45. In other words, the accuracy in identifying hate speech in the zero-shot setting is 63 per cent. The model has an F1 score of 81 per cent in this setting, with a notably high recall of 87 per cent as well.

@@

The results of the one-shot learning experiments are presented in Table \@ref(tab:oneshot-summary). Tables \@ref(tab:oneshot-racism), \@ref(tab:oneshot-sexism), and \@ref(tab:oneshot-hate), in Appendix \@ref(appendixboneshot) provide additional detail. The model has 37 matches and 23 mismatches in the racist category, and 36 matches and 24 mismatches in the sexist category. In contrast with the result generated from zero-shot learning, the model performs approximately the same in identifying racist comments and identifying sexist comments. The overall ratio of matches and mismatches is 73:47. In other words, the accuracy of identifying hate speech in the one-shot setting is 61 per cent. The general performance in the one-shot setting is worse than in the zero-shot setting, with lower accuracy, recall, precision, and F1 score overall.

@@

The results of the single-category, few-shot learning, experiments are presented in Table \@ref(tab:fewshotsingle-summary). Tables \@ref(tab:fewshotsingle-racism), \@ref(tab:fewshotsingle-sexism), \@ref(tab:fewshotsingle-hate), in Appendix \@ref(appendixbfewshotsingle) provide additional detail. The model has 41 matches and 19 mismatches in the racist category, and 42 matches and 18 mismatches in the sexist category. Like the one-shot setting, the model performs almost equally well at identifying racist comments compared with identifying sexist comments. The general performance in the single-category few-shot learning setting is slightly more accurate than performance in other settings, with an accuracy of 69 per cent compared with 61 per cent in the one-shot setting and 63 per cent in the zero-shot setting. However, the model in this context has the lowest F1 score of the settings considered thus far at 61 per cent.

@@

The results of the mixed-category few-shot experiments are presented in Table \@ref(tab:fewshotmixed-summary). Tables \@ref(tab:fewshotmixed-racism) and Table \@ref(tab:fewshotmixed-sexism) in Appendix \@ref(appendxbmuxedubsnorinstruction) provide additional detail. Among the ten sets of examples, Example Set 1 yields the best performance in terms of accuracy (91 per cent) and F1 score (87 per cent) for racist comments and Example Set 10 yields the best performance by the same metrics for sexist comments (accuracy of 89 per cent, F1 score of 85 per cent). The example set that yields the worst results in identifying racist text in terms of F1 score is Example Set 8, which has an F1 score of 70 per cent (and the lowest accuracy at 73 per cent) for this dataset. The example set that yields the worst results in identifying sexist text in terms of F1 score is Example Set 9, which has an F1 score of 69 per cent (and the lowest accuracy at 74 per cent) for this dataset. The differences between Example Sets 1, 8, 9, and 10 suggest that, although the models are provided with same number of examples, the content of the examples also affects how the model makes inferences. Overall, the mixed-category few-shot setting performs approximately the same at identifying sexist text and racist text. It is also has higher accuracy and F1 score than the zero-shot, one-shot, and single-category few-shot settings for both racist and sexist text.

@@

The unique generated answers are listed in Table \@ref(tab:fewshotmixedanswersnoinstruct). These are the response of GPT-3 that we obtain when we ask the model to classify statements, but do not provide examples that would serve to limit the responses. Under the mixed-category setting, the model is observed to generate answers that are out of scope. For instance, other than 'sexist', 'racist', and 'neither', we also see answers such as 'transphobic', 'hypocritical', 'Islamophobic', and 'ableist'. In some cases, the model even classifies a text passage into more than one category, such as 'sexist, racist' and 'sexist, misogynistic'. The full list contains eighty different answers instead of three.

The results presented for each category of text include the classifications of comments that were labelled as 'neither' and the category in question. For the purposes of our analysis, a classification was considered a true positive if the answer outputted by GPT-3 contained a category that matched the comment's label. For example, if a comment was labelled 'sexist' and the comment was classified by the model as 'sexist, racist', this was considered a true positive in the classification of sexist comments. If a comment was labelled 'sexist' and the comment was classified by the model as 'racist', 'transphobic', 'neither', etc. this was considered a false negative.

@@

The results of the mixed-category few-shot learning, with instruction, experiments are presented in Tables \@ref(tab:fewshotmixedinstruct-matrix) and \@ref(tab:fewshotmixedinstruct-summary). And Tables \@ref(tab:fewshotmixedinstruct-racism), and \@ref(tab:fewshotmixedinstruct-sexism), in Appendix \@ref(appendxbmuxedubstryctub) provide additional detail. With the addition of an instruction in the prompt, Example Set 10 remains the best performing example set in terms of accuracy (89 per cent) and F1 score (84 per cent) for sexist text, however Example Set 4 now performs best for racist text in terms of both accuracy (89 per cent) and F1 score (85 per cent). When considering the classification of racist and sexist speech overall, the models perform similarly with and without instruction when classifying racist text, but the model appears to perform slightly better at identifying sexist text when the instruction is omitted.

@@

knitr::kable(table,
             caption="Performance of mixed-category few-shot learning in text classification, with instruction",
            col.names = c("Example set", "Category", "Accuracy (%)", "Precision (%)", "Recall (%)", "F1 (%)"),
             booktabs=TRUE,
             linesep = "") %>%
  kableExtra::row_spec(seq(2, 20, by=2), hline_after = TRUE)
```

@@

```{r zeroshot-racism, message=FALSE}
zero_shot_result_racism %>%
  knitr::kable(caption = "Classification of racist statements with zero-shot learning",
               booktabs = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "GPT-3 classification" = 2)) %>% 
  kableExtra::kable_styling(latex_options = "hold_position")
```

```{r zeroshot-sexism}
zero_shot_result_sexism %>%
  knitr::kable(caption = "Classification of sexist statements with zero-shot learning",
               booktabs = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "GPT-3 classification"=2)) %>% 
  kableExtra::kable_styling(latex_options = "hold_position")
```

```{r zeroshot-hate}
zero_shot_result_hate %>%
  knitr::kable(caption = "Classification of hate speech with zero-shot learning",
               booktabs = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "GPT-3 classification"=2)) %>% 
  kableExtra::kable_styling(latex_options = "hold_position")
```

>>>>

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: true
    toc: false
    fig_caption: true
    latex_engine: xelatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
  -  \usepackage{amsmath}
biblio-style: apalike
title: "Detecting Hate Speech with GPT-3"
thanks: "Code and data are available at: https://github.com/kelichiu/GPT3-hate-speech-detection. We gratefully acknowledge the support of Gillian Hadfield, the Schwartz Reisman Institute for Technology and Society, and OpenAI for providing access to GPT-3 under the academic access program. We thank two anonymous reviews and the editor, as well as Amy Farrow, Haoluan Chen, John Giorgi, Mauricio Vargas Sepúlveda, Monica Alexander, Noam Kolt, and Tom Davidson for helpful discussions and suggestions. Please note that we have added asterisks to racial slurs and other offensive content in this paper, however the inputs and outputs did not have these.  Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Ke-Li Chiu
  affiliation: University of Toronto
- name: Annie Collins
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: University of Toronto and Schwartz Reisman Institute
abstract: "Sophisticated language models such as OpenAI's GPT-3 can generate hateful text that targets marginalized groups. Given this capacity, we are interested in whether large language models can be used to identify hate speech and classify text as sexist or racist. We use GPT-3 to identify sexist and racist text passages with zero-, one-, and few-shot learning. We find that with zero- and one-shot learning, GPT-3 can identify sexist or racist text with an accuracy between 57 per cent and 68 per cent depending on the category of text and type of learning. With few-shot learning, the model's accuracy can be as high as 88 per cent. Large language models have a role to play in hate speech detection, and with further development could eventually be used to counter hate speech."
keywords: "GPT-3; natural language processing; quantitative analysis; hate speech."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
endnote: no
graphics: yes
---

@@

The text produced by sophisticated NLP models can be hateful. In particular, there have been many examples of text being generated that targets marginalized groups based on their sex, race, sexual orientation, and other characteristics. For instance, 'Tay' was a Twitter chatbot released by Microsoft in 2016. Within hours of being released, some of its tweets were sexist. Large language models are trained on enormous datasets from various, but primarily internet-based, sources. This means they usually contain untruthful statements, human biases, and abusive language. Even though models do not possess intent, they do produce text that is offensive or discriminatory, and thus cause unpleasant, or even triggering, interactions [@bender2021dangers].

Often the datasets that underpin these models consist of, essentially, the whole public internet. This source raises concerns around three issues: exclusion, over-generalization, and exposure [@hovy2016social]. Exclusion happens due to the demographic bias in the dataset. In the case of language models that are trained on US and UK English scraped from the Internet, datasets may be disproportionately white, male, and young. Therefore, it is not surprising to see white supremacist, misogynistic, and ageist content being over-represented in training datasets [@bender2021dangers]. Over-generalization stems from the assumption that what we see in the dataset represents what occurs in the real world. Words such as 'always', 'never', 'everybody', or 'nobody' are frequently used for rhetorical purpose instead of their literal meanings. But NLP models do not always recognize this and make inference based on generalized statements using these words. For instance, hate speech commonly uses generalized language for targeting a group such as 'all' and 'every', and a model trained on these statements may generate similarly generalized and harmful statements. Finally, exposure refers to the relative attention, and hence consideration of importance, given to something. In the context of NLP this may be reflected in the emphasis on English-language created under particular circumstances, rather than another language or circumstances that may be more prevalent.

@@

Even though large NLP models do not have a real understanding of language, the vocabularies and the construction patterns of hateful language can be thought of as known to them. We show that this knowledge can be used to identify abusive language and even hate speech. We consider 120 different extracts that have been categorized as 'racist', 'sexist', or 'neither' in single-category settings (zero-shot, one-shot, and few-shot) and 243 different extracts in mixed-category few-shot settings. We ask GPT-3 to classify these based on zero-, one-, and few-shot learning, with and without instruction. We find that the model performs best with mixed-category few-shot learning. In that setting the model can accurately classify around 86 per cent of the racist extracts and 88 per cent of sexist extracts. If language models can be used to identify abusive language, then not only is there potential for them to counter the production of abusive language by humans, but they could also potentially self-police.

The remainder of this paper is structured as follows: Section \@ref(background) provides background information about language models and GPT-3 in particular. Section \@ref(methods) introduces our dataset and our experimental approach to zero-, one-, and few-shot learning. Section \@ref(results) conveys the main findings of those experiments. And Section \@ref(discussion) adds context and discusses some implications, next steps, and weaknesses. Appendices \@ref(appendxa), \@ref(appendxb), and \@ref(appendxc) contain additional information.

@@

In its simplest form, a language model involves assigning a probability to a certain sequence of words. For instance, the sequence 'the cat in the hat' is probably more likely than 'the cat in the computer'. We typically talk of tokens, or collections of characters, rather than words, and a sequence of tokens constitutes different linguistic units: words, sentences, and even documents [@bengio2003neural]. Language models predict the next token based on inputs. If we consider each token in a vocabulary as a dimension, then the dimensionality of language quickly becomes large [@rosenfeld2000two]. Over time a variety of statistical language models have been created to nonetheless enable prediction. The n-gram is one of the earliest language models. It works by considering the co-occurrence of tokens in a sequence. For instance, given the four-word sequence, 'the cat in the', it is more likely that the fifth word is 'hat' rather than 'computer'. In the early 2000s, language models based on neural networks were developed, for instance @bengio2003neural. These were then built on by word embeddings language models in the 2010s in which the distance between tokens represents how related those tokens are, for instance @turian2010word. In 2017, @vaswani2017attention introduced the Transformer, which marked a new era for language models. The Transformer is a network architecture for neural networks that can be trained more quickly than many other approaches [@vaswani2017attention]. Now most representative pre-trained language models, such as Google's Bidirectional Encoder Representations from Transformers (BERT) [@devlin2018bert], as well as OpenAI's Generative Pre-trained Transformer (GPT)-2 [@radford2019language], and GPT-3 [@brown2020language], are built on this architecture. These models are widely used, for instance BERT is used by Google search.

@@

There is no commonly held definition of hate speech. Different legal jurisdictions have different definitions, as do different companies and other groups. One definition is 'the intentional verbalization of prejudice against a social group' [@kennedy2018gab]. Detecting hate speech is difficult because the definition of hate speech varies depending on a complex intersection of the topic of the assertion, the context, the timing, outside events, and the identity of speaker and recipient [@schmidt2017survey]. Moreover, it is difficult to distinguish hate speech from offensive language [@davidson2017automated]. Hate speech detection is of interest to academic researchers in a variety of domains including computer science [@addressinghatespeech] and sociology [@davidson2017automated]. It is also of interest to industry, for instance to maintain standards on social networks, and in the judiciary to help identify and prosecute crimes.  Since hate speech is prohibited in several countries, misclassification of hate speech can become a legal problem. For instance, in Canada, speech that contains 'public incitement of hatred' or 'wilful promotion of hatred' is specified by the Criminal Code [@act2021justice]. Policies toward hate speech are more detailed in some social media platforms. For instance, the Twitter Hateful Conduct Policy states:

@@

There has been a large amount of research focused on detecting hate speech. As part of this process, various hate speech datasets have been created and examined. For instance, @waseem2016hateful detail a dataset that captures hate speech in the form of racist and sexist language that includes domain expert annotation. They use Twitter data, and annotate 16,914 tweets: 3,383 as sexist, 1,972 as racist, and 11,559 as neither. They had a high degree of annotator agreement. Most of the disagreements were to do with sexism, and often explained by an annotator lacking apparent context. @davidson2017automated train a classifier to distinguish between hate speech and offensive language. To define hate speech, they use an online 'hate speech lexicon containing words and phrases identified by internet users as hate speech'. Even these datasets have bias. For instance, @davidson2019racial found racial bias in five different sets of Twitter data annotated for hate speech and abusive language. They found that tweets written in African American English are more likely to be labeled as abusive.

@@

We use the ETHOS dataset of @mollas2020ethos. ETHOS is based on comments from YouTube and Reddit. The ETHOS YouTube data is collected through Hatebusters [@anagnostou2018hatebusters]. Hatebusters is a platform that collects comments from YouTube and assigns a 'hate' score to them using a support vector machine. That hate score is only used to decide whether to consider the comment further or not. The Reddit data is collected from the Public Reddit Data Repository [@baumgartner2020pushshift]. The classification is done by contributors to a crowd-sourcing platform. They are first asked whether an example contains hate speech, and then, if it does, whether it incites violence and other additional details. The dataset has two variants: binary and multi-label. In the binary dataset comments are classified as hate or non-hate based. In the multi-label variant comments are evaluated on measures that include violence, gender, race, ability, religion, and sexual orientation. The dataset that we use is as provided by the ETHOS dataset and hence contain typos, misspelling, and offensive content.

We begin with all of the 998 statements in the ETHOS dataset that have a binary classification of hate speech or not hate speech. Of these, the 433 statements that contain hate speech additionally have labels that classify the content. For instance, does the comment have to do with violence, gender, race, nationality, disability, etc. We initially considered all of the 136 statements that contain race-based hate speech, but we focus on the 76 whose race-based score is at least 0.5, meaning that at least 50 per cent of annotators agreed. Similarly, we initially considered all of the 174 statements that contain gender-based hate speech, and again focused on the 84 whose gender-based score is at least 0.5. To create a balanced dataset, we select 120 of these statements to result in 60 sexist statements and 60 racist statements. For comparison purposes, we randomly draw 120 statements that are classified as not hate speech. The balanced sample enables us to more easily understand our results.

@@

Under the few-shot setting, the model is given three examples. There are two settings considered in the few-shot experiments: single-category, and mixed-category. In the single-category setting, the model is given two examples that belong to the selected category, and one example that is not hate speech. The answer is expected to be binary. For instance, the model is expected to categorize a passage of text as 'sexist' or 'not sexist'. An example of this is:

@@

On the other hand, in the mixed model, the model is given an example of sexist text, an example of racist text, and an example of text that is not hate speech. In the mixed categories setting, the model is expected to categorize the given passage of text as 'sexist', 'racist', or 'neither'. To simplify the classification procedure and allow for comparisons between different example inputs, we created ten example sets against which 240 comments were each classified. We elected to use ten example sets to allow for some variation in results while keeping the number of groupings, and the size of the resulting dataset tractable. The example sets each consist of one sexist, one racist, and one 'neither' comment and were created using the comments at indexes one through ten from each dataset. Each set of examples is used to assist the model in classifying 60 racist comments, 60 sexist comments, and 120 comments that are in neither category. The inputted comments were the first _n_ + 1 comments from their respective datasets with the _i_^th^ comment dropped corresponding to the use of example set _i_. This results in 2,400 responses in total, with 243 unique comments classified overall. The ten sets of examples are shown in Appendix \@ref(appendxa). An example of this is:

@@

We assess GPT-3's performance in all settings using accuracy, precision, recall, and F1 score. Accuracy is the proportion of correctly classified comments (hate speech and non-hate speech) out of all comments classified. Precision is the proportion of hate speech comments correctly classified out of all comments classified as hate speech (both correctly and incorrectly). Recall is the proportion of hate speech comments correctly classified out of all hate speech comments in the dataset (both correctly and incorrectly classified). The F1 score is the harmonic mean of precision and recall. In the case of hate speech classification, we see it as better to have a model with high recall, meaning a model that can identify a relatively high proportion of the hate speech text within a dataset. But the F1 score can provide a more well-rounded metric for model performance and comparison.

@@

The overall results of the zero-shot experiments are presented in Table \@ref(tab:zeroshot-summary), and Appendix \@ref(appendixbzeroshot) provides additional detail. The model has 34 matches (true positives and negatives) and 26 mismatches (false positives and negatives) in the sexist category, and 41 matches and 19 mismatches in the racist category. The model performs more accurately when identifying racist comments, with an accuracy of 68 per cent, compared with identifying sexist comments, with an accuracy of 57 per cent, but the F1 score for each type of hate speech is roughly the same. The overall ratio of matches and mismatches is 75:45. In other words, the accuracy in identifying hate speech in the zero-shot setting is 63 per cent. The model has an F1 score of 81 per cent in this setting, with a notably high recall of 87 per cent.

@@

The results of the one-shot learning experiments are presented in Table \@ref(tab:oneshot-summary), and Appendix \@ref(appendixboneshot) provides additional detail. The model has 37 matches and 23 mismatches in the racist category, and 36 matches and 24 mismatches in the sexist category. In contrast with the result generated from zero-shot learning, the model performs roughly the same in terms of identifying racist comments and identifying sexist comments. The overall ratio of matches and mismatches is 73:47. In other words, the accuracy of identifying hate speech in the one-shot setting is 61 per cent. The general performance in the one-shot setting is worse than in the zero-shot setting, with lower accuracy, recall, precision, and F1 score overall.

@@

The results of the single-category, few-shot learning, experiments are presented in Table \@ref(tab:fewshotsingle-summary), and Appendix \@ref(appendixbfewshotsingle) provides additional detail. The model has 41 matches and 19 mismatches in the racist category, and 42 matches and 18 mismatches in the sexist category. Like the one-shot setting, the model performs almost equally well in terms of identifying racist comments compared with identifying sexist comments. The general performance in the single-category few-shot learning setting is slightly more accurate than performance in other settings, with an accuracy of 69 per cent compared with 61 per cent in the one-shot setting and 63 per cent in the zero-shot setting. But, in this context we see the lowest F1 score of the settings considered thus far at 61 per cent.

@@

The results of the mixed-category few-shot experiments are presented in Table \@ref(tab:fewshotmixed-summary), and Appendix \@ref(appendxbmuxedubsnorinstruction) provides additional detail. Among the ten sets of examples, Example Set 1 yields the best performance in terms of accuracy (91 per cent) and F1 score (87 per cent) for racist comments and Example Set 10 yields the best performance by the same metrics for sexist comments (accuracy of 89 per cent, F1 score of 85 per cent). The example set that yields the worst results in identifying racist text in terms of F1 score is Example Set 8, which has an F1 score of 70 per cent (and the lowest accuracy at 73 per cent) for this dataset. The example set that yields the worst results in identifying sexist text in terms of F1 score is Example Set 9, which has an F1 score of 69 per cent (and the lowest accuracy at 74 per cent) for this dataset. The differences between Example Sets 1, 8, 9, and 10 suggest that, although the models are provided with the same number of examples, the content of the examples also affects how the model makes inferences. Overall, the mixed-category few-shot setting performs roughly the same in terms of identifying sexist text and racist text. It is also has higher accuracy and F1 score than the zero-shot, one-shot, and single-category few-shot settings for both racist and sexist text.

@@

The unique generated answers are listed in Table \@ref(tab:fewshotmixedanswersnoinstruct). These are the response of GPT-3 that we obtain when we ask the model to classify statements, but do not provide examples that would serve to limit the responses. Under the mixed-category setting, the model is generates many answers that are out of scope. For instance, other than 'sexist', 'racist', and 'neither', we also see answers such as 'transphobic', 'hypocritical', 'Islamophobic', and 'ableist'. In some cases, the model even classifies a text passage into more than one category, such as 'sexist, racist' and 'sexist, misogynistic'. The full list contains eighty different answers instead of three.

The results presented for each category of text include the classifications of comments that were labelled as 'neither' and the category in question. For the purposes of our analysis, a classification was considered a true positive if the answer outputted by GPT-3 contained a category that matched the comment's label. For example, if a comment was labelled 'sexist' and the comment was classified by the model as 'sexist, racist', this was considered a true positive in the classification of sexist comments. If a comment was labelled 'sexist' and the comment was classified by the model as 'racist', 'transphobic', 'neither', etc, then this was considered a false negative.

@@

The results of the mixed-category few-shot learning, with instruction, experiments are presented in Tables \@ref(tab:fewshotmixedinstruct-matrix) and \@ref(tab:fewshotmixedinstruct-summary), and Appendix \@ref(appendxbmuxedubstryctub) provides additional detail. With the addition of an instruction in the prompt, Example Set 10 remains the best performing example set in terms of accuracy (89 per cent) and F1 score (84 per cent) for sexist text, however Example Set 4 now performs best for racist text in terms of both accuracy (89 per cent) and F1 score (85 per cent). When considering the classification of racist and sexist speech overall, the models perform similarly with and without instruction when classifying racist text, but the model appears to perform slightly better at identifying sexist text when the instruction is omitted.

@@

knitr::kable(table,
             caption="Performance of mixed-category few-shot learning in text classification, with instruction",
            col.names = c("Example set", "Category", "Accuracy (%)", "Precision (%)", "Recall (%)", "F1 (%)"),
             booktabs=TRUE,
            digits = 0,
             linesep = "") %>%
  kableExtra::row_spec(seq(2, 20, by=2), hline_after = TRUE)
```

@@

```{r zeroshot-racism, message=FALSE}
zero_shot_result_racism %>%
  knitr::kable(caption = "Classification of racist statements with zero-shot learning",
               booktabs = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "GPT-3 classification" = 2)) %>% 
  kableExtra::kable_styling(font_size = 8) %>%
  kableExtra::kable_styling(latex_options = "hold_position") 
```

```{r zeroshot-sexism}
zero_shot_result_sexism %>%
  knitr::kable(caption = "Classification of sexist statements with zero-shot learning",
               booktabs = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "GPT-3 classification"=2)) %>% 
  kableExtra::kable_styling(font_size = 8) %>%
  kableExtra::kable_styling(latex_options = "hold_position")
```
